---
title: "HW2"
author: "Xiangxiong Feng"
date: "2023-10-02"
output: github_document
---


# Problem 1

```{r}
library(tidyverse)
library(lubridate)
library(dplyr)
```
# a. clean the data in pols-month.csv

```{r}
polsmonth_df = 
  read_csv('data/pols-month.csv') |>
  janitor:: clean_names() |>
  separate(mon, into = c('year', 'month'), sep = 4) |>
  separate(month, into = c('month', 'day'), sep = 3) |>
  mutate(
    month = replace(month, month == '-01', 'Jan'),
    month = replace(month, month == '-02', 'Feb'),
    month = replace(month, month == '-03', 'Mar'),
    month = replace(month, month == '-04', 'Apr'),
    month = replace(month, month == '-05', 'May'),
    month = replace(month, month == '-06', 'Jun'),
    month = replace(month, month == '-07', 'Jul'),
    month = replace(month, month == '-08', 'Aug'),
    month = replace(month, month == '-09', 'Sep'),
    month = replace(month, month == '-10', 'Oct'),
    month = replace(month, month == '-11', 'Noc'),
    month = replace(month, month == '-12', 'Dec'),
    month = str_to_upper(month)
  ) |>
  rename(president = prez_dem) |>
  mutate(
    president = replace(president, president == '0', 'gop'),
    president = replace(president, president == '1', '0'),
    president = replace(president, president == 'gop', '1')
    ) |>
  select(-prez_gop,-day)
  
  
```

# b. clean data in snp.csv

```{r}
snp_df = 
  read_csv('data/snp.csv') |>
  janitor::clean_names() |>
  mutate(
    date = parse_date_time(date, order = c('mdy'))) |>
  separate(date, into = c('year', 'month'), sep = 4) |>
  separate(month, into = c('month', 'day'), sep = 3) |>
  mutate(
    month = replace(month, month == '-01', 'Jan'),
    month = replace(month, month == '-02', 'Feb'),
    month = replace(month, month == '-03', 'Mar'),
    month = replace(month, month == '-04', 'Apr'),
    month = replace(month, month == '-05', 'May'),
    month = replace(month, month == '-06', 'Jun'),
    month = replace(month, month == '-07', 'Jul'),
    month = replace(month, month == '-08', 'Aug'),
    month = replace(month, month == '-09', 'Sep'),
    month = replace(month, month == '-10', 'Oct'),
    month = replace(month, month == '-11', 'Noc'),
    month = replace(month, month == '-12', 'Dec'),
    month = str_to_upper(month)) |>
  select(-day)
  



```


# c. clean data in unemployment.csv

```{r}
unemployment_df = 
  read_csv('data/unemployment.csv') |>
  janitor::clean_names() |>
  pivot_longer(
    jan:dec,
    names_to = 'month',
    values_to = 'unemployment_rate'
  ) |>
  mutate(
    month = str_to_upper(month),
  
  )

```
# d. Joining dataset

```{r}
pol_snp_df = 
  left_join(polsmonth_df,snp_df ) |>
  mutate(year = as.numeric(year))

final_df =
  left_join(pol_snp_df, unemployment_df)
```

The final dataset includes 11 variables, some important variables are year, month
and unemployment rate. The range of the year is from 1947 to 2015.





# Problem 2 

# a. import and clean Mr. Trash Wheel

```{r}
library( readxl)
Mr.Trash_wheel_df = 
  read_excel('data/202309 Trash Wheel Collection Data.xlsx', sheet = 'Mr. Trash Wheel') |>
  janitor::clean_names() |>
  drop_na(dumpster) |>
  mutate(
    homes_powered = (weight_tons*500)/30,
    Wheel_Type = 'Mr.Trash Wheel',
    year = as.numeric(year)
  )



```

# b. import and clean Professor Trash Wheel

```{r}
Professor_Trash_df = 
  read_excel('data/202309 Trash Wheel Collection Data.xlsx', sheet = 'Professor Trash Wheel') |>
  janitor::clean_names() |>
  drop_na(dumpster) |>
  mutate(
    homes_powered = (weight_tons*500)/30,
    Wheel_Type = 'Professor Trash Wheel'
  )

```


# c. Import and clean Gwynnda Trash Wheel

```{r}
Gwynnda_Trash_df = 
  read_excel('data/202309 Trash Wheel Collection Data.xlsx', sheet = 'Gwynnda Trash Wheel') |>
  janitor::clean_names() |>
  drop_na(dumpster) |>
  mutate(
    homes_powered = (weight_tons*500)/30,
    Wheel_Type = 'Gwynnda Trash Wheel'
  )

Number_cigarette =
  filter(Gwynnda_Trash_df, year==2021, month == 'July') 


```

# d. Combine dataset to produce a single tidydataset

```{r}
Final_TrashWheel_df =
  bind_rows(Mr.Trash_wheel_df, Professor_Trash_df, Gwynnda_Trash_df) |>
  select(Wheel_Type, everything()) 
  


  
```

# e. Description:

There are 845 observation in the resulting dataset. The dataset has `r ncol(Final_TrashWheel_df)` columes and `r nrow(Final_TrashWheel_df)` rows.Some key variables includes Wheel_Type, which shows which type the trash wheel is; Homes Powered, which is calculate from Weight and average household energy usage; data, which 
shows data information. By the available data, the total weight of trash collected by 
Professor Trash Wheel is `r sum(Professor_Trash_df$weight_tons)`. The total number of cigarette butts collected by Gwynnda in July of 2021 is `r sum(Number_cigarette$cigarette_butts)`.



# Problem 3


# a. import and clean the dataset of baseline

```{r}
baseline_df =
  read_csv('data/MCI_baseline.csv', skip = 1) |>
  janitor::clean_names() |>
  mutate(
    sex = 
      case_match(
        sex,
        1 ~ 'male',
        0 ~ 'female'),
    sex = as.factor(sex),
    apoe4 =
      case_match(
        apoe4,
        1 ~ 'carrier',
        0 ~ 'noncarrier' ),
    age_at_onset = as.numeric(age_at_onset))

baseline_filtered_df = 
  filter(baseline_df, age_at_onset != '.')
```
# proportion of women in the study are APOE4 carriers
```{r}
#women in the study are APOE4 carriers
sex_apoe = 
  read_csv('data/MCI_baseline.csv', skip = 1) |>
  janitor::clean_names() |>
  mutate(
    sex = 
      case_match(
        sex,
        1 ~ 'male',
        0 ~ 'female'),
    sex = as.factor(sex),
    apoe4 =
      case_match(
        apoe4,
        1 ~ 'carrier',
        0 ~ 'noncarrier' ),
  age_at_onset = as.numeric(age_at_onset)) |>
  select(sex,apoe4) |>
  filter(sex == 'female') 

```

There are 483 obs and 6 variables in the resulting data set. During the import process, the value of variables: sex and apoe4 are converted to non-numerical value. There are 97 obs develop
MCI. The average current age is `r mean(baseline_df$current_age)`. `r (length(which(sex_apoe$apoe4 == 'carrier')))/211` of women in the study are APOE4 carriers.


# b. import and clean the dataset of longitudinally observed biomarker values

```{r}
amyloid_df =
  read.csv('data/mci_amyloid.csv', skip = 1) |>
  janitor::clean_names() |>
  rename(id = study_id)



 #tidy the dataset 
amyloid_tydi_df = 
  pivot_longer(amyloid_df,
    baseline:time_8,
    names_to = 'time_period',
    values_to = 'observed_biomarker_value') 
 

```

The resulting data has 2435 obs and 3 variables, which are study_id, time_period
and observed_biomarker_value. The data shows the biomarker values for each participants 
in different time periods. 

# c. description and calculation 
```{r EVA = FALSE}

unqiue_baseline = setdiff(baseline_df$id, amyloid_df$id)
length(unqiue_baseline)

unqiue_amyloid = setdiff(amyloid_df$id, baseline_df$id)
length(unqiue_amyloid)

baseline_amyloid_df = 
  inner_join(baseline_filtered_df, amyloid_tydi_df, by ='id')
write.csv(baseline_amyloid_df, file = 'data/resluting_data_Problem 3.csv')
```

There are 8 participants only appear in  baseline and 12 only appear in amyloid dataset.
There are total 470 participants who appear in both datasets are retained. The resulting dataset has 470 obs and 8 variables. The dataset shows current age, age 
at onset and biomarker value for each participant in 5 different time periods.


