---
title: "HW2"
author: "Xiangxiong Feng"
date: "2023-10-02"
output: github_document
---


# Problem 1

```{r}
library(tidyverse)
```
# a. clean the data in pols-month.csv

```{r}
polsmonth_df = 
  read_csv('data/pols-month.csv') |>
  janitor:: clean_names() |>
  separate(mon, into = c('year', 'month'), sep = 4) |>
  separate(month, into = c('month', 'day'), sep = 3) |>
  mutate(
    month = replace(month, month == '-01', 'Jan'),
    month = replace(month, month == '-02', 'Feb'),
    month = replace(month, month == '-03', 'Mar'),
    month = replace(month, month == '-04', 'Apr'),
    month = replace(month, month == '-05', 'May'),
    month = replace(month, month == '-06', 'Jun'),
    month = replace(month, month == '-07', 'Jul'),
    month = replace(month, month == '-08', 'Aug'),
    month = replace(month, month == '-09', 'Sep'),
    month = replace(month, month == '-10', 'Oct'),
    month = replace(month, month == '-11', 'Noc'),
    month = replace(month, month == '-12', 'Dec'),
    month = str_to_upper(month)
  ) |>
  rename(president = prez_dem) |>
  mutate(
    president = replace(president, president == '0', 'gop'),
    president = replace(president, president == '1', '0'),
    president = replace(president, president == 'gop', '1')
    ) |>
  select(-prez_gop,-day)
  
  
```

# b. clean data in snp.csv

```{r}
library(lubridate)
library(dplyr)
snp_df = 
  read_csv('data/snp.csv') |>
  janitor::clean_names() |>
  mutate(
    date = parse_date_time(date, order = c('mdy'))) |>
  separate(date, into = c('year', 'month'), sep = 4) |>
  separate(month, into = c('month', 'day'), sep = 3) |>
  mutate(
    month = replace(month, month == '-01', 'Jan'),
    month = replace(month, month == '-02', 'Feb'),
    month = replace(month, month == '-03', 'Mar'),
    month = replace(month, month == '-04', 'Apr'),
    month = replace(month, month == '-05', 'May'),
    month = replace(month, month == '-06', 'Jun'),
    month = replace(month, month == '-07', 'Jul'),
    month = replace(month, month == '-08', 'Aug'),
    month = replace(month, month == '-09', 'Sep'),
    month = replace(month, month == '-10', 'Oct'),
    month = replace(month, month == '-11', 'Noc'),
    month = replace(month, month == '-12', 'Dec'),
    month = str_to_upper(month)) |>
  select(-day)
  



```


# c. clean data in unemployment.csv

```{r}
unemployment_df = 
  read_csv('data/unemployment.csv') |>
  janitor::clean_names() |>
  pivot_longer(
    jan:dec,
    names_to = 'month',
    values_to = 'unemployment_rate'
  ) |>
  mutate(
    month = str_to_upper(month),
  
  )

```
# d. Joining dataset

```{r}
pol_snp_df = 
  left_join(polsmonth_df,snp_df ) |>
  mutate(year = as.numeric(year))

final_df =
  left_join(pol_snp_df, unemployment_df)
```

The final dataset includes 11 variables, some important variables are year, month
and unemployment rate. The range of the year is from 1947 to 2015.





# Problem 2 

# a. import and clean Mr. Trash Wheel

```{r}
library( readxl)
Mr.Trash_wheel_df = 
  read_excel('data/202309 Trash Wheel Collection Data.xlsx', sheet = 'Mr. Trash Wheel') |>
  janitor::clean_names() |>
  drop_na(dumpster) |>
  mutate(
    homes_powered = (weight_tons*500)/30,
    Wheel_Type = 'Mr.Trash Wheel',
    year = as.numeric(year)
  )



```

# b. import and clean Professor Trash Wheel

```{r}
Professor_Trash_df = 
  read_excel('data/202309 Trash Wheel Collection Data.xlsx', sheet = 'Professor Trash Wheel') |>
  janitor::clean_names() |>
  drop_na(dumpster) |>
  mutate(
    homes_powered = (weight_tons*500)/30,
    Wheel_Type = 'Professor Trash Wheel'
  )

```


# c. Import and clean Gwynnda Trash Wheel

```{r}
Gwynnda_Trash_df = 
  read_excel('data/202309 Trash Wheel Collection Data.xlsx', sheet = 'Gwynnda Trash Wheel') |>
  janitor::clean_names() |>
  drop_na(dumpster) |>
  mutate(
    homes_powered = (weight_tons*500)/30,
    Wheel_Type = 'Gwynnda Trash Wheel'
  )

Number_cigarette =
  filter(Gwynnda_Trash_df, year==2021, month == 'July') 


```

# d. Combine dataset to produce a single tidydataset

```{r}
Final_TrashWheel_df =
  bind_rows(Mr.Trash_wheel_df, Professor_Trash_df, Gwynnda_Trash_df) |>
  select(Wheel_Type, everything()) 
  


  
```

# e. Description:

There are 845 observation in the resulting dataset. Some key variables includes Wheel_Type, which shows which type the trash wheel is; Homes Powered, which is 
calculate from Weight and average household energy usage; data, which shows data information. By the available data, the total weight of trash collected by 
Professor Trash Wheel is `r sum(Mr.Trash_wheel_df$weight_tons)`. The total number of cigarette butts collected by Gwynnda in July of 2021 is `r sum(Number_cigarette$cigarette_butts)`.



# Problem 3


# a.

```{r}
baseline_df =
  read_csv('data/MCI_baseline.csv', skip = 1) |>
  janitor::clean_names() |>
  mutate(
    sex = 
      case_match(
        sex,
        1 ~ 'male',
        0 ~ 'female'),
    sex = as.factor(sex),
    apoe4 =
      case_match(
        apoe4,
        1 ~ 'carrier',
        0 ~ 'noncarrier' ))
  
  

  

  
  
```





